---
title: "usfertilizer"

output: rmarkdown::github_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

![Travis-CI Build Status](https://travis-ci.org/wenlong-liu/usfertilizer.svg?branch=master)
[![AppVeyor Build Status](https://ci.appveyor.com/api/projects/status/github/wenlong-liu/usfertilizer?branch=master&svg=true)](https://ci.appveyor.com/project/wenlong-liu/usfertilizer)


## County-lelel nutrients data from 1945 to 2012
Usfertilizer summarized the estimated county level data from USGS and provided a clean version using Tidyverse. 

Please note that USGS does not endorse this package and this package does not include manure data. Also data from 1986 is not available for now.

## Installment
For now, you can install it via my Github Repository.

```{r, eval=FALSE}
# install.package("devtools")   #In case you have not installed it.
devtools::install_github("wenlong-liu/usfertilizer")
```

In the future, I will try to submit it for CRAN.  You can also install it via:
```{r, eval= FALSE}
install.packages("usfertilizer")
# not available yet.
```

#  Get started
## Import data and related libraries

```{r, message=FALSE, warning=FALSE}
require(usfertilizer)
require(tidyverse)
data("us_fertilizer_county")
```

## Summary of the dataset

The dataset, named by us_fertilizer_county, contains `r length(us_fertilizer_county$FIPS)` observations and 11 variables.  Details are available by using `?us_fertilizer_county`.

```{r}
glimpse(us_fertilizer_county)
```

# Examples
## Example 1: Find out the top 10 counties with most nitrogen appliation in 2008.

```{r}
# plot the top 10 nitrogen application in year 2008.
# Reorder to make the plot more cleanner.
year_plot = 2008
us_fertilizer_county %>%
  filter(Fertilizer == "N" & Year == year_plot) %>%
  top_n(10, Quantity) %>%
  ggplot(aes(x=reorder(paste(County,State, sep = ","), Quantity), Quantity, fill = Quantity))+
  scale_fill_gradient(low = "blue", high = "darkblue")+
  geom_col()+
  ggtitle(paste("Top 10 counties with most fertilizer application in the year of", year_plot)) + 
  scale_y_continuous(name = "Nitrogen from commecial fertilization (kg)")+
  scale_x_discrete(name = "Counties")+
  coord_flip()+
  theme_bw()
```

## Example 2: Find out the top 10 states with most nitrogen appliation in 1980.

```{r}
# plot the top 10 states with P application in year 1980.
# Reorder to make the plot more cleanner.
year_plot = 1980
us_fertilizer_county %>%
  filter(Fertilizer == "P" & Year == 1980) %>% 
  group_by(State) %>% 
  summarise(p_application = sum(Quantity)) %>% 
  as.data.frame() %>% 
  top_n(10, p_application) %>%
  ggplot(aes(x=reorder(State, p_application), p_application))+
  scale_fill_gradient(low = "blue", high = "darkblue")+
  geom_col()+
  ggtitle(paste("Top 10 States with most Phosphrus application in the year of", year_plot)) + 
  scale_y_continuous(name = "Phosphrus from commecial fertilization (kg)")+
  scale_x_discrete(name = "States")+
  theme_bw()+
  coord_flip()
  
```


## Example 3: Plot the N and P input into farms for NC and SC from 1945 to 2010
```{r, message=F, warning=F}
year_plot = seq(1945, 2010, 1)
states = c("NC","SC")

us_fertilizer_county %>% 
  filter(State %in% states & Year %in% year_plot &
           Farm.Type == "farm") %>% 
  group_by(State, Year, Fertilizer) %>% 
  summarise(Quantity = sum(Quantity)) %>% 
  ggplot(aes(x = as.numeric(Year), y = Quantity, color=State)) +
  geom_point() +
  geom_line()+
  scale_x_continuous(name = "Year")+
  scale_y_continuous(name = "Nutrient input quantity (kg)")+
  facet_wrap(~Fertilizer, scales = "free", ncol = 2)+
  ggtitle("Estimated nutrient inputs into arable lands by commercial fertilizer\nfrom 1945 to 2010 in Carolinas")+
  theme_bw()

```

# Comments and Questions.

If you have any problems or questions, feel free to open an issue [here](https://github.com/wenlong-liu/usfertilizer/issues).

# Lisence
GPL







